---
title: "Mediation Under Confounding on Real Dataset"
author: "Yuhan Pu"
date: "2024-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import package
These packages would probably be used in the future
```{r cars}
library (nlme)
library (lme4)
library(lmerTest)
library(mediation)
library(car)
library(GGally)
library(ggplot2)
library(dplyr)
library(purrr)
```

Import the Dataset

```{r pressure, echo=FALSE}
data <- read.csv('All_data_S.csv')
data <- subset(data,Strategy!=2)
```

Define Each Variables

```{r}
sub_data=subset(data,Completion==0)
age=sub_data$age
bmi=sub_data$BMI
str=sub_data$Strategy
td=sub_data$TD
subject=sub_data$subject##be used as the random effect#
```

```{r}

```



Models Among Each Variables (Using random effect)

```{r}
mod_td<-lmer(td~age+str+bmi+(1|subject),data=sub_data)
mod_bmi<-lmer(bmi~age+(1|subject),data=sub_data)
mod_str<-glmer(as.factor(str)~bmi+age+(1|subject),data=sub_data,family = binomial(link = 'logit'))
```

```{r}
dat = data.frame(td,age,str,bmi)
```



Calculating TE: TE=E[Y(1)-Y(0)]

```{r}
dat_te_1 = dat
dat_te_0 = dat
dat_te_1$bmi = quantile(dat$bmi,0.975)
dat_te_0$bmi = quantile(dat$bmi,0.025)

dat_te_me = dat[c('bmi','age')]
dat_te_me$bmi = quantile(dat$bmi,0.975)
predicted_probs <- predict(mod_str, dat_te_me, type='response')
n <- length(predicted_probs)
dat_te_1$str <- rbinom(n=n, prob=predicted_probs, size=1)
dat_te_me$bmi = quantile(dat$bmi,0.025)
predicted_probs <- predict(mod_str, dat_te_me, type='response')
n <- length(predicted_probs)
dat_te_0$str <- rbinom(n=n, prob=predicted_probs, size=1)

TE = mean(predict(mod_td,dat_te_1)-predict(mod_td,dat_te_0))
```

Calculating CDE: CDE(m)=E[Y(1,m)-Y(0,m)]

```{r}
dat_cde_1 = dat
dat_cde_0 = dat
dat_cde_1$bmi = quantile(dat$bmi,0.975)
dat_cde_0$bmi = quantile(dat$bmi,0.025)
dat_cde_1$str = 1
dat_cde_0$str = 1
CDE = mean(predict(mod_td,dat_cde_1)-predict(mod_td,dat_cde_0))
```

Calculating NDE: NDE=E[Y(1,M(0))-Y(0,M(0))]

```{r}
dat_nde_1 = dat
dat_nde_0 = dat
dat_nde_1$bmi = quantile(dat$bmi,0.975)
dat_nde_0$bmi = quantile(dat$bmi,0.025)
dat_nde_me = dat[c('bmi','age')]
dat_nde_me$bmi = quantile(dat$bmi,0.025)
dat_nde_1$str = rbinom(n=n,prob = predict(mod_str,dat_nde_me,type='response'),size=1)
dat_nde_0$str = rbinom(n=n,prob = predict(mod_str,dat_nde_me,type='response'),size=1)
NDE = mean(predict(mod_td,dat_nde_1)-predict(mod_td,dat_nde_0))
```

Calculating NIE: NIE=E[Y(0,M(1))-Y(0,M(0))]

```{r}
dat_nie_1 = dat
dat_nie_0 = dat
dat_nie_1$bmi = quantile(dat$bmi,0.025)
dat_nie_0$bmi = quantile(dat$bmi,0.025)

dat_nie_me = dat[c('bmi','age')]
dat_nie_me$bmi = quantile(dat$bmi,0.975)
dat_nie_1$str = rbinom(n=n,prob = predict(mod_str,dat_nie_me,type='response'),size=1)

dat_nie_me$bmi = quantile(dat$bmi,0.025)
dat_nie_0$str = rbinom(n=n,prob = predict(mod_str,dat_nie_me,type='response'),size=1)

NIE = mean(predict(mod_td,dat_nie_1)-predict(mod_td,dat_nie_0))
```

Print the Result

```{r}
cat("TE: ",TE, "\nCDE: ",CDE,"\nNDE: ",NDE, "\nNIE: ", NIE)
```
By this we noticed that NIE here to small so we need to fit a regression with str~bmi to check its statistical significance.
```{r}
a_model<-glmer(str~bmi+(1|subject), data=sub_data,family = binomial(link = 'logit'))
summary(a_model)
```

It shows significance even though its coefficient is too small. So we still want to combine strategy=0 and 1 together, and strategy=2 isolated.


Then we tried that again with combining strategy 0 and 1 together and 2 isolated and do the whole process again
```{r}
data <- read.csv('All_data_S.csv')
data$Strategy <- ifelse(data$Strategy == 2, 1, 0)
```

```{r}
sub_data=subset(data,Completion==0)
age=sub_data$age
bmi=sub_data$BMI
str=sub_data$Strategy
td=sub_data$TD
subject=sub_data$subject##be used as the random effect#
mod_td<-lmer(td~age+str+bmi+(1|subject),data=sub_data)
mod_bmi<-lmer(bmi~age+(1|subject),data=sub_data)
mod_str<-glmer(as.factor(str)~bmi+age+(1|subject),data=sub_data,family = binomial(link = 'logit'))
dat = data.frame(td,age,str,bmi)
dat_te_1 = dat
dat_te_0 = dat
dat_te_1$bmi = quantile(dat$bmi,0.975)
dat_te_0$bmi = quantile(dat$bmi,0.025)

dat_te_me = dat[c('bmi','age')]
dat_te_me$bmi = quantile(dat$bmi,0.975)
predicted_probs <- predict(mod_str, dat_te_me, type='response')
n <- length(predicted_probs)
dat_te_1$str <- rbinom(n=n, prob=predicted_probs, size=1)
dat_te_me$bmi = quantile(dat$bmi,0.025)
predicted_probs <- predict(mod_str, dat_te_me, type='response')
n <- length(predicted_probs)
dat_te_0$str <- rbinom(n=n, prob=predicted_probs, size=1)

TE = mean(predict(mod_td,dat_te_1)-predict(mod_td,dat_te_0))
dat_cde_1 = dat
dat_cde_0 = dat
dat_cde_1$bmi = quantile(dat$bmi,0.975)
dat_cde_0$bmi = quantile(dat$bmi,0.025)
dat_cde_1$str = 1
dat_cde_0$str = 1
CDE = mean(predict(mod_td,dat_cde_1)-predict(mod_td,dat_cde_0))
dat_nde_1 = dat
dat_nde_0 = dat
dat_nde_1$bmi = quantile(dat$bmi,0.975)
dat_nde_0$bmi = quantile(dat$bmi,0.025)
dat_nde_me = dat[c('bmi','age')]
dat_nde_me$bmi = quantile(dat$bmi,0.025)
dat_nde_1$str = rbinom(n=n,prob = predict(mod_str,dat_nde_me,type='response'),size=1)
dat_nde_0$str = rbinom(n=n,prob = predict(mod_str,dat_nde_me,type='response'),size=1)
NDE = mean(predict(mod_td,dat_nde_1)-predict(mod_td,dat_nde_0))
dat_nie_1 = dat
dat_nie_0 = dat
dat_nie_1$bmi = quantile(dat$bmi,0.025)
dat_nie_0$bmi = quantile(dat$bmi,0.025)

dat_nie_me = dat[c('bmi','age')]
dat_nie_me$bmi = quantile(dat$bmi,0.975)
dat_nie_1$str = rbinom(n=n,prob = predict(mod_str,dat_nie_me,type='response'),size=1)

dat_nie_me$bmi = quantile(dat$bmi,0.025)
dat_nie_0$str = rbinom(n=n,prob = predict(mod_str,dat_nie_me,type='response'),size=1)

NIE = mean(predict(mod_td,dat_nie_1)-predict(mod_td,dat_nie_0))
cat("TE: ",TE, "\nCDE: ",CDE,"\nNDE: ",NDE, "\nNIE: ", NIE)

```
```{r}
b_model<-lmer(td~bmi+str+(1|subject), data=sub_data)
summary(b_model)
```

